---
title: "Reproducible Research: Peer Assessment 1"
author: "Jeremy Peters"
date: "Nov 4, 2017"
output: 
  html_document:
    keep_md: true
---

```{r "setup", include=FALSE}
require("knitr")
require("lattice")
#opts_knit$set(root.dir = "C:/JP Docs/Data Science Certification/WD/")
```

## Loading and preprocessing the data
### Read  the CSV  file in table format and create a data frame 
### Convert the date values to a  R date object
```{r readdata, echo  =  TRUE}
df <- read.csv("activity.csv", header = TRUE)
df$date <- as.Date(df$date)
```


## What is mean total number of steps taken per day?
### Use aggregate function to calculate the total number of steps per day
### Create  a histogram for total number of steps taken each day


```{r histogramTotalSteps1,  echo  =  TRUE}
totalSteps <- aggregate(steps ~ date,data = df, sum, na.rm = TRUE)

hist(totalSteps$steps,col = "blue", xlab = "No. Steps", main = "Histogram Total Number of Steps Taken Each Day")

```

### Determine mean and median total number of steps taken per day
```{r meanmediansteps,  echo  =  TRUE}

meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medianSteps <- median(totalSteps$steps, na.rm = TRUE)
meanSteps
medianSteps
```
### The mean  total number of steps taken per day is 10766.
### The median total number of steps taken per day is 10765.


## What is the average daily activity pattern?
###Create a time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r activitypatternplot, fig.height=4,fig.width=8, echo  =  TRUE}

avgStepsPerInt <- with(na.omit(df),tapply(steps,interval,mean))

plot(names(avgStepsPerInt), avgStepsPerInt, type = "l", xlab = "5-minute Interval", 
    ylab = "Number of Steps", main = "Average Daily Activity Pattern")


```

###Determine the 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps
```{r maxinterval,  echo  =  TRUE}
maxIndex <- which(avgStepsPerInt == max(avgStepsPerInt))
maxInterval <- df$interval[maxIndex]
maxInterval
```
###The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of step is 835.

## Imputing missing values

### Calculate  the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r numberMissingValues, echo  =  TRUE}
mv <- sum(is.na(df))
mv
```
### The total number of missing values in the dataset: 2304

### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be 
### sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

### The following strategy includes:
### 1 Create a copy of the dataset
### 2 Loop through each record in the copy of the data set
### 3 Check if there is a missing value in the steps field
### 4.Call a function to calculate the mean steps per interval if a steps value is missing
### 5.Update the the missing value with the calculated mean steps per interval in the copy of the data set 


### Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r createNewData,echo  =  TRUE}

df2<-aggregate(steps~interval,data=df,mean,na.rm=TRUE)
getMeanSteps<-function(interval){
    df2[df2$interval==interval,]$steps
}

df3<-df   
count=0           
for(i in 1:nrow(df3)){
    if(is.na(df3[i,]$steps)){
        df3[i,]$steps<-getMeanSteps(df3[i,]$interval)
        count=count+1
    }
}



```
### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number ### of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the 
### impact of imputing missing data on the estimates of the total daily number of steps?
```{r histogramTotalSteps2,echo  =  TRUE}
totalSteps2 <- aggregate(steps ~ date,data = df3, sum)

hist(totalSteps2$steps,col = "blue", xlab = "Number of Steps", main = "Histogram Total Number of Steps Taken Each Day")
meanSteps2 <- mean(totalSteps2$steps)
medianSteps2 <- median(totalSteps2$steps)
meanSteps2
medianSteps2
```


### The mean  total number of steps taken per day is 10766.
### The median total number of steps taken per day is 10766.

### There is almost no impact on the mean and median values by  imputing missing data on the estimates of the total daily  
### number of steps. The mean value remains the same at 10766. The median value changed from 10765 to 10766. The frequency ### has increased as can be seen in particular the center bar of the histogram.

## Are there differences in activity patterns between weekdays and weekends?
### For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for ### this part.
### 1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r createFactorVariable, echo  =  TRUE}

wdays <- weekdays(df3$date)
satsun <- (wdays == "Saturday" | wdays == "Sunday")
dfactor <- factor(satsun, labels = list("weekday", "weekend"))
df3$dtype <- dfactor
```

### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average 
### number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub 
### repository to see an example of what this plot should look like using simulated data.

```{r timeSeriesPanelPlot, echo  =  TRUE}
totalSteps3 <- aggregate(df3$steps, list(DayType = df3$dtype, Interval = df3$interval), mean)

xyplot(totalSteps3$x ~ totalSteps3$Interval | totalSteps3$DayType, layout = c(1, 
    2), xlab = "Interval", ylab = "Number of Steps", type = "l")
```
