---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r "setup", include=FALSE}
require("knitr")

opts_knit$set(root.dir = "C:/JP Docs/Data Science Certification/WD/")
```

## Loading and preprocessing the data
### Read  the CSV  file in table format and create a data frame 
### Convert the data to a  R date object
```{r readdata, echo  =  TRUE}
df <- read.csv("activity.csv", header = TRUE)
df$date <- as.Date(df$date)
```


## What is mean total number of steps taken per day?
### Create  a histogram for total number of steps taken each day
```{r histogram,  echo  =  TRUE}
totalSteps <- with(df,tapply(df$steps,df$date,FUN = sum, na.rm = TRUE))

hist(totalSteps, xlab = "Total Steps", ylab = "Number of Days", main = "Total Number of Steps Taken Each Day", col = "green")
qplot(totalSteps, geom = "histogram", binwidth = 1000, xlab = "Total Steps", ylab = "Number of Days", main = "Total Number of Steps Taken Each Day", fill = ..count..)

```

### Determine mean and median total number of steps taken per day
```{r meanmediansteps,  echo  =  TRUE}

meanSteps <- mean(totalSteps, na.rm = TRUE)
medianSteps <- median(totalSteps, na.rm = TRUE)
meanSteps
medianSteps
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
